\thetitle

### 2008-06-29 letter.tex
Generate tarball; finalized sections in references.

\name{%
Stefan~Karpinski\\
Elizabeth~M.~Belding\\
Kevin~C.~Almeroth\\
John~R.~Gilbert%
}

@@

% It was mentioned that it was unclear how the transformations allow to understand the effect and drawbacks of each simplified uniform, marginal or conditional model.
In the revised paper, we have observed that the uniform, marginal and conditional transformations are all multiplications by specific low-rank matrices.
Since low-rank matrices are by definition uninvertible, the information content of the original traffic matrix is irrevocably destroyed by these transformations.
The revised paper discusses this issue in two places:
\begin{enumerate}
\item Introduction, page 2, column 2, paragraph 2;
\item Discussion, page 15, column 2, paragraph 1.
\end{enumerate}
In the discussion section, moreover, we propose matrix factorization, rather matrix multiplication as a less destructive and thus more faithful method of reducing the complexity, size and dimension of traffic representations.

@@

% TODO: address complexity in discussion.
This issue is now addressed in our discussion section.
We admit that large matrix representations are more computationally costly than simpler models.
The purpose of this work, however, is to establish how traffic can be represented and generated realistically at all, not how it can be done most efficiently.
Currently, realistic traffic modeling is an unsolved problem.
We observe, however, that generating application traffic for a simulation takes orders of magnitude less \caps{CPU}-time than running the simulation itself.
Therefore, traffic model generation complexity is considered a non-issue.

@@

We already briefly addressed this issue in the original version of the paper:
we have evaluated a variety of other performance metrics, but omit them here due to lack of space.
Moreover, the performance comparisons for the metrics presented here are indicative of other performance metrics.
We address this now in our discussion section in greater detail.

@@

\begin{reviewer}{1}
Does 24-hour trace data used in the paper really good enough?
\end{reviewer}
The adequacy of the 24-hour trace is discussed in the discussion section: in short, 24 hours of data is not enough to draw general conclusions about all traffic, but is enough to demonstrate the potential effectiveness of \caps{GMM}, and provide counter-examples to the effectiveness of uniform, marginal and conditional modeling techniques;
this is all we attempt to do here.
Further validation of methods on larger sets of data is required in the future.

\begin{reviewer}{1}
What theoretical and computational tools of modern linear algebra can be adopted and used for traffic generation specifically? One thing they mentioned as their current work is using matrix factorization and clustering technique to extract hidden structure from a large and diverse body of real traffic. But this haven’t verified yet and isn’t it too soon to claim this?
\end{reviewer}
We provide a concrete example of a powerful technique that can be applied immediately to the \caps{GMM}: we use nonnegative matrix factorization to understand and explain flow packet behaviors.
This technique and some exemplary results from its application are shown in the discussion section.
It is also noted that all the traffic models used in this paper are written in only 19 lines of Matlab code, demonstrating the power of and simplicity of linear representation and the \caps{GMM} for traffic modeling explorations.

\begin{reviewer}{1}
Wouldn‘t it be better if the authors come up with the criteria to determine whether the seven properties characterize the traced traffic adequately instead of proving the adequacy of GMM model by showing the six performance metrics closely resembled those of traced traffic? For example, by comparing distribution of packet size of GMM model and that of traced data.
\end{reviewer}
The seven properties used in the \caps{GMM} are not meant to be comprehensive, but rather representative.
In the section in which we introduce the metrics, we state that we have collected and analyzed other performance metrics in our, experiments, but these performance metrics are representative of the overall results.
At a higher level, comparing empirical distributions of models to the originals is meaningless:
if the model is based on the trace, and includes packet size distribution then the two will necessarily match;
running experiments is not necessary to show that.
The real question is whether this matters or not:
does it affect experimental results?
We refer the reviewer to our earlier work for a more in-depth discussion of the issue~\cite{Karpinski07:realism,Karpinski07:cbr-failure}.

\begin{reviewer}{1}
Paxson and Floyd observed TCP traffic which protocol itself has closed-loop behavior. However, can we generalized this to the entire network as the authors mentioned?
\end{reviewer}
We note in our revised related work section that the work of \FHC~\cite{Hernandez06:dissertation} perfectly complements ours in that it provides the means to emulate the closed-loop behavior of applications using \caps{TCP}, but does not provide any models of application behavior.
Our work, on the other hand, provides the necessary models of applications behavior.
Taken together, the \caps{TCP} flow generation technique of \FHC and our application-level workload modeling address both halves of the closed-loop network traffic modeling problem.
See Related Work, page 3, column 2, paragraph 2 for our discussion of this issue.

@@

\closing{Yours Respectfully,}

>>>>

@@

% It was mentioned that it was unclear how the transformations allow to understand the effect and drawbacks of each simplified uniform, marginal or conditional model.
In the revised paper, we have observed that the uniform, marginal and conditional transformations are all multiplications by specific low-rank matrices.
Since low-rank matrices are by definition uninvertible, the information content of the original traffic matrix is irrevocably destroyed by these transformations.
The revised paper discusses this issue in Section 1 (p. 2, col. 2, par. 2), Section 4.4, and Section 7 (p. 2, col. 2, par. 2) sections.
In the discussion section (Section 7), moreover, we propose matrix factorization, rather matrix multiplication as a less destructive and thus more faithful method of reducing the complexity, size and dimension of traffic representations.

@@

This issue is now addressed in our discussion: Section 7.2.
We admit that large matrix representations are more computationally costly than simpler models.
The purpose of this work, however, is to establish how traffic can be represented and generated realistically at all, not how it can be done most efficiently.
Currently, realistic traffic modeling is an unsolved problem.
We observe, however, that generating application traffic for a simulation takes orders of magnitude less \caps{CPU}-time than running the simulation itself.
Therefore, traffic model generation complexity is considered a non-issue.

@@

We already briefly addressed this issue in the original version of the paper:
we have evaluated a variety of other performance metrics, but omit them here due to lack of space.
Moreover, the performance comparisons for the metrics presented here are indicative of other performance metrics.
We address this now in our discussion section in greater detail: see Section 7.1.

@@

\begin{reviewer}{1}
Does 24-hour trace data used in the paper really good enough?
\end{reviewer}
The adequacy of the 24-hour trace is discussed in the discussion section: in short, 24 hours of data is not enough to draw general conclusions about all traffic, but is enough to demonstrate the potential effectiveness of \caps{GMM}, and provide counter-examples to the effectiveness of uniform, marginal and conditional modeling techniques;
this is all we attempt to do here.
Further validation of methods on larger sets of data is required in the future.
We now address this issue in the discussion: see Section 7.1.

\begin{reviewer}{1}
What theoretical and computational tools of modern linear algebra can be adopted and used for traffic generation specifically? One thing they mentioned as their current work is using matrix factorization and clustering technique to extract hidden structure from a large and diverse body of real traffic. But this haven’t verified yet and isn’t it too soon to claim this?
\end{reviewer}
We provide a concrete example of a powerful technique that can be applied immediately to the \caps{GMM}: we use nonnegative matrix factorization to understand and explain flow packet behaviors.
This technique and some exemplary results from its application are shown in the discussion section (Section 7.3).
It is also noted that all the traffic models used in this paper are written in only 19 lines of Matlab code, demonstrating the power of and simplicity of linear representation and the \caps{GMM} for traffic modeling explorations.

\begin{reviewer}{1}
Wouldn‘t it be better if the authors come up with the criteria to determine whether the seven properties characterize the traced traffic adequately instead of proving the adequacy of GMM model by showing the six performance metrics closely resembled those of traced traffic? For example, by comparing distribution of packet size of GMM model and that of traced data.
\end{reviewer}
The seven properties used in the \caps{GMM} are not meant to be comprehensive, but rather representative.
In Section 5.4, we state that we have collected and analyzed other performance metrics in our, experiments, but these performance metrics are representative of the overall results.
We now further address this issue in Section 7.1 as well.
To address the reviewer's specific point, at a higher level, comparing empirical distributions of models to the originals is meaningless:
if the model is based on the trace, and includes packet size distribution then the two will necessarily match;
running experiments is not necessary to show that.
The real question is whether this matters or not:
does it affect experimental results?
We refer the reviewer to our earlier work for a more in-depth discussion of the issue~\cite{Karpinski07:realism,Karpinski07:cbr-failure}.

\begin{reviewer}{1}
Paxson and Floyd observed TCP traffic which protocol itself has closed-loop behavior. However, can we generalized this to the entire network as the authors mentioned?
\end{reviewer}
We note in our revised related work section that the work of \FHC~\cite{Hernandez06:dissertation} perfectly complements ours in that it provides the means to emulate the closed-loop behavior of applications using \caps{TCP}, but does not provide any models of application behavior.
Our work, on the other hand, provides the necessary models of applications behavior.
Taken together, the \caps{TCP} flow generation technique of \FHC and our application-level workload modeling address both halves of the closed-loop network traffic modeling problem.
See Section 2 for our discussion of this issue.

@@

Yours Respectfully,

\begin{quote}
Stefan~Karpinski\\
Elizabeth~M.~Belding\\
Kevin~C.~Almeroth\\
John~R.~Gilbert
\end{quote}
